---
title: "Reading Hillary Clinton's Emails"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Why?

I like politics, am interested in NLP, and would like to know what Hillary and her 
co-workers were emailing about that caused such a fuss.

## Some info about the data

* There are just under 8,000 emails in this data set <br>
* Kaggle performed some processing to extract and clean the emails<br>
* The code that prepared the data is available at <a href="https://github.com/benhamner/hillary-clinton-emails" target="_blank">this GitHub Repo</a>

```{r check out emails, echo=FALSE}
library(dplyr)
library(stringr)
library(tidytext)
library(ggplot2)
library(scales)
library(wordcloud)
library(viridis)
library(tidyr)

my.colors2 <- viridis(5, begin = 0, end = 0.75, option = "A" )
my.colors3 <- viridis(10, begin = 0, end = 0.85, option = "D" )

email.csv <- read.csv(file = "./data/Emails.csv", header = TRUE, nrows = 7946, 
                      stringsAsFactors = FALSE)
email <- tbl_df(email.csv)
rm(email.csv)
glimpse(email)
```

## What words show up most often in the email subjects

I used the MetadataSubject Variable for this plot:
```{r subject word count plot}
tidy.email.subject <- email %>%
                unnest_tokens(word, MetadataSubject)

data("stop_words")
tidy.email.subject <- tidy.email.subject %>%
                anti_join(stop_words)

tidy.email.subject %>%  
    count(word, sort = TRUE) %>%  
    filter(n > 50) %>%  
    mutate(word = reorder(word, n)) %>%  
    ggplot(aes(word, n)) +  
    geom_bar(stat = "identity") +  
    xlab(NULL) +  
    coord_flip()

tidy.email.subject %>%
    mutate(word = str_extract(word, "[a-z']+")) %>%
    count(word) %>%
    with(wordcloud(word, n, scale=c(13, .5), max.words = 50, random.color = TRUE, colors=my.colors2))
```

## Haiti vs. Iran
* Each word is rated as either positive or negative for this examination using bing taxonomy

The prior plot showed that Haiti and Iran were the two countries that showed 
up most frequently in email's subjects

So let's compare the words in those emails:

```{r echo=FALSE}
## Let's compare words in emails with subject "haiti" and "iran"
haiti.email <- email %>% 
    filter(grepl("haiti", MetadataSubject, ignore.case = TRUE)) %>%
    unnest_tokens(word, ExtractedBodyText) %>%
    anti_join(stop_words)

iran.email <- email %>% 
    filter(grepl("iran", MetadataSubject, ignore.case = TRUE))  %>%
    unnest_tokens(word, ExtractedBodyText) %>%
    anti_join(stop_words)

count(haiti.email, word, sort=TRUE)
count(iran.email, word, sort=TRUE)

haiti.percent <- haiti.email %>%  
    mutate(word = str_extract(word, "[a-z']+")) %>%
    count(word) %>%  
    transmute(word, haiti = n / sum(n))

iran.percent <- iran.email %>%  
    mutate(word = str_extract(word, "[a-z']+")) %>%
    count(word) %>%  
    transmute(word, iran = n / sum(n))

frequency <- iran.percent %>%
                left_join(haiti.percent) %>%
                ungroup()

ggplot(frequency, aes(x = iran, y = haiti, color = abs(iran - haiti))) +
    geom_abline(color = "gray40", lty = 2) +
    geom_jitter(alpha = 0.1, size = 2.5, width = 0.3, height = 0.3) +
    geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
    scale_x_log10(labels = percent_format()) +
    scale_y_log10(labels = percent_format()) +
    scale_color_gradient(limits = c(0, 0.001), 
                         low = "darkslategray4", high = "gray75") +
    ##facet_wrap(~author, ncol = 2) +
    theme(legend.position="none") +
    labs(y = "Haiti", x = "Iran")
```

## Everybody loves a wordcloud - More on Haiti vs. Iran

```{r}

iran.email %>%
    mutate(word = str_extract(word, "[a-z']+")) %>%
    count(word) %>%
    with(wordcloud(word, n, scale=c(75, .5) ,max.words = 50, random.color = TRUE, colors=my.colors2))

haiti.email %>%
    mutate(word = str_extract(word, "[a-z']+")) %>%
    count(word) %>%
    with(wordcloud(word, n, scale=c(13, .5) ,max.words = 50, random.color = TRUE, colors=my.colors3))
```

## Sentiment Analysis of Haiti and Iran emails

```{r}
haiti.email.sentiment <- email %>% 
    filter(grepl("haiti", MetadataSubject, ignore.case = TRUE)) %>%
    group_by(Id) %>%
    mutate(index = row_number()) %>%
    ungroup() %>%
    unnest_tokens(word, ExtractedBodyText) 

bingPositive <- get_sentiments("bing") %>%
    filter(sentiment == "positive")

haiti.email.sentiment2 <- haiti.email.sentiment %>%
    inner_join(get_sentiments("bing")) %>%
    count(Id, sentiment) %>%
    spread(sentiment, n, fill=0) %>%
    mutate(sentiment = positive - negative)

haiti.index <- 1:52
df.haiti <- as.data.frame(haiti.email.sentiment2)
df.haiti <- cbind(df.haiti, haiti.index)
haiti.email.sentiment2 <- tbl_df(df.haiti)

ggplot(haiti.email.sentiment2, aes(haiti.index, sentiment)) +
    geom_col(show.legend = TRUE, color = "green") 
    
iran.email.sentiment <- email %>% 
    filter(grepl("iran", MetadataSubject, ignore.case = TRUE)) %>%
    group_by(Id) %>%
    mutate(index = row_number()) %>%
    ungroup() %>%
    unnest_tokens(word, ExtractedBodyText) 

bingPositive <- get_sentiments("bing") %>%
    filter(sentiment == "positive")

iran.email.sentiment2 <- iran.email.sentiment %>%
    inner_join(get_sentiments("bing")) %>%
    count(Id, sentiment) %>%
    spread(sentiment, n, fill=0) %>%
    mutate(sentiment = positive - negative)

iran.index <- 1:30
df.iran <- as.data.frame(iran.email.sentiment2)
df.iran <- cbind(df.iran, iran.index)
iran.email.sentiment2 <- tbl_df(df.iran)

ggplot(iran.email.sentiment2, aes(iran.index, sentiment)) +
    geom_col(show.legend = TRUE, color = "red")
```

## Comparing the taxonomies
* Afinn and Bing and NRC use different methods. Let's see how they compare:

```{r}
iran.afinn <- iran.email.sentiment %>%
    inner_join(get_sentiments("afinn")) %>%
    group_by(Id) %>%
    summarise(sentiment = sum(score)) %>%
    mutate(method = "AFINN")

iran.afinn <- iran.afinn %>% tibble::rownames_to_column(var = "index1")
    
iran.bing.and.nrc <- bind_rows(iran.email.sentiment %>%
                                   inner_join(get_sentiments("bing")) %>%
                                   mutate(method = "Bing"),
                               iran.email.sentiment %>%
                                   inner_join(get_sentiments("nrc") %>%
                                                  filter(sentiment %in% c("positive", 
                                                                          "negative"))) %>%
                                   mutate(method = "NRC")) %>%
                    count(method, index = Id, sentiment) %>%
                        spread(sentiment, n, fill = 0) %>%
                        mutate(sentiment = positive - negative)

iran.bing.and.nrc <- iran.bing.and.nrc %>% tibble::rownames_to_column(var = "index1")
                    
bind_rows(iran.afinn, iran.bing.and.nrc) %>%
    ggplot(aes(index1, sentiment, fill = method)) +
    geom_col(show.legend = FALSE) +
    facet_wrap(~method, ncol = 1, scales = "free_y")
```



